### 2019-07-12 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
OH LA BOULETTE : en anglais, « repartition » n'existe pas. On dit « distribution » ou « allocation »
- https://www.wordreference.com/enfr/repartition
- https://www.deepl.com/translator#en/fr/repartition

We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of the repartition of static devices $(S_k)_k$, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead. This is not possible in IoT contexts as several completely independent networks may operate in a single place.} is possible.
We want to find the stationary repartition of devices into channels that maximizes the probability of having a successful transmission.

@@

Note that solving the optimization problem provides the optimal \emph{real values} $(D_k^*)_k$, which have to be rounded to find the optimal \emph{integer} number of devices for channel $k$.
Any rounding choice will give about the same repartition, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
%
$\widehat{D_k} = \lfloor D_k^* \rfloor$ for $1 \leq k < K$, and $\widehat{D_{K}} = D - \sum_{k=1}^{K - 1} \widehat{D_k}$.

@@

% Simulation parameters
We suppose a network with $S + D = 2000$ end-devices, and one IoT gateway.
Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\eg, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
The RF band is divided in $K = 10$ channels.
Each static device only uses one channel, and their uneven repartition in the $10$ channels is chosen as $(S_1,\cdots, S_{K}) = S \times (0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases. The dynamic devices have access to all the channels, and use learning algorithms.
%to find the least loaded.
%
We simulate the network during $10^6$ discrete time slots, during which each device transmits on average $1000$ packets (\ie, the learning time is about $1000$ steps, for each algorithm).
We tried similar experiments with other values for $K$ and this repartition vector, and results were similar for non-homogeneous repartitions.

Clearly, the problem is less interesting for a homogeneous repartition, as all channels are equivalent for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[scale=0.90]{perf_learning.eps}
    \caption[Learning with \UCB{} and Thomson Sampling, with many dynamic devices]{Learning with \textcolor{red}{\UCB} and \textcolor{gold}{Thomson Sampling}, with many dynamic devices. The learning gain, for each device, decreases with the proportion of dynamic devices in the network. Note that the values ($16\%$ etc) depend on the repartition of static devices into the $K$ channels (\ie, $S_1,\dots,S_K$) but the general profile of this plot does \emph{not} depend much on these parameters.}
    \label{fig:41:perf_learning}
\end{figure}

@@

\begin{figure}[!h]
    \centering
    \subfloat[][10\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_10.eps}
    \label{fig:41:ch2_10}}
    %
    \subfloat[][30\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_30.eps}
    \label{fig:41:ch2_30}}
    %
    \vspace*{-10pt}
    \subfloat[][50\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_50.eps}
    \label{fig:41:ch2_50}}
    %
    \subfloat[][100\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_100.eps}
    \label{fig:41:ch2_100}}
    \caption[Performance of the \UCB bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.]{Performance of the \textcolor{blue}{\UCB} bandit algorithm for the special case of uniform repartition of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.}
    \label{fig:41:figure4appendix}
\end{figure}

@@

\paragraph{A safety check.}
%
We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform repartition of static devices (\ie, $\forall k, S_k = S/K$), to check that learning approach (here, only \UCB)
also gives interesting gain of performance, and achieve a close-to optimal successful transmission rate.
Except for the limit case of $100\%$ of dynamic devices, which corresponds to Figures~\ref{fig:41:100intelligent} and \ref{fig:41:ch2_100}, where the uniform access performs as well as the optimal oracle solution, the MAB-based approach almost instantly outperforms the baseline.

>>>>

We investigate here the optimal policy that can be achieved if the dynamic devices have a perfect knowledge of the distribution of static devices $(S_k)_k$, and a fully centralized decision making\footnote{~This optimal policy needs an \emph{oracle} seeing the entire system, and affecting all the dynamic devices, once and for all, for a fixed stationary scenario, in order to avoid any signaling overhead. This is not possible in IoT contexts as several completely independent networks may operate in a single place.} is possible.
We want to find the stationary distribution of devices into channels that maximizes the probability of having a successful transmission.

@@

Note that solving the optimization problem provides the optimal \emph{real values} $(D_k^*)_k$, which have to be rounded to find the optimal \emph{integer} number of devices for channel $k$.
Any rounding choice will give about the same distribution, up-to a difference of only one device by channel, and so we chose to round from below for the first channels:
%
$\widehat{D_k} = \lfloor D_k^* \rfloor$ for $1 \leq k < K$, and $\widehat{D_{K}} = D - \sum_{k=1}^{K - 1} \widehat{D_k}$.

@@

% Simulation parameters
We suppose a network with $S + D = 2000$ end-devices, and one IoT gateway.
Each device sends packets following a Bernoulli process, of probability $p = 10^{-3}$ (\eg, this is realistic: one packet sent about every $20$ minutes, for time slots of $1\mathrm{s}$).
The RF band is divided in $K = 10$ channels.
Each static device only uses one channel, and their uneven distribution in the $10$ channels is chosen as $(S_1,\cdots, S_{K}) = S \times (0.3, \, 0.2, \, 0.1, \, 0.1, \, 0.05, \, 0.05, \, 0.02, \, 0.08, \, 0.01,$ $0.09)$, to keep the same proportions when $S$ decreases. The dynamic devices have access to all the channels, and use learning algorithms.
%to find the least loaded.
%
We simulate the network during $10^6$ discrete time slots, during which each device transmits on average $1000$ packets (\ie, the learning time is about $1000$ steps, for each algorithm).
We tried similar experiments with other values for $K$ and this distribution vector, and results were similar for non-homogeneous repartitions.

Clearly, the problem is less interesting for a homogeneous distribution, as all channels are equivalent for dynamic devices, and so even with $D$ small in comparison to $S$, the system behaves like in Figure~\ref{fig:41:100intelligent}, where the performance of the five approaches are very close.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[scale=0.90]{perf_learning.eps}
    \caption[Learning with \UCB{} and Thomson Sampling, with many dynamic devices]{Learning with \textcolor{red}{\UCB} and \textcolor{gold}{Thomson Sampling}, with many dynamic devices. The learning gain, for each device, decreases with the proportion of dynamic devices in the network. Note that the values ($16\%$ etc) depend on the distribution of static devices into the $K$ channels (\ie, $S_1,\dots,S_K$) but the general profile of this plot does \emph{not} depend much on these parameters.}
    \label{fig:41:perf_learning}
\end{figure}

@@

\begin{figure}[!h]
    \centering
    \subfloat[][10\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_10.eps}
    \label{fig:41:ch2_10}}
    %
    \subfloat[][30\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_30.eps}
    \label{fig:41:ch2_30}}
    %
    \vspace*{-10pt}
    \subfloat[][50\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_50.eps}
    \label{fig:41:ch2_50}}
    %
    \subfloat[][100\% of intelligent devices]{\includegraphics[width=0.52\textwidth]{ch2_100.eps}
    \label{fig:41:ch2_100}}
    \caption[Performance of the \UCB bandit algorithm for the special case of uniform distribution of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.]{Performance of the \textcolor{blue}{\UCB} bandit algorithm for the special case of uniform distribution of the static devices, when the proportion of intelligent devices in the network increases, from $10\%$ to $100\%$.}
    \label{fig:41:figure4appendix}
\end{figure}

@@

\paragraph{A safety check.}
%
We include another simulation in Figure~\ref{fig:41:figure4appendix}, with a uniform distribution of static devices (\ie, $\forall k, S_k = S/K$), to check that learning approach (here, only \UCB)
also gives interesting gain of performance, and achieve a close-to optimal successful transmission rate.
Except for the limit case of $100\%$ of dynamic devices, which corresponds to Figures~\ref{fig:41:100intelligent} and \ref{fig:41:ch2_100}, where the uniform access performs as well as the optimal oracle solution, the MAB-based approach almost instantly outperforms the baseline.

### 2019-07-12 2-Chapters/4-Chapter/IEEE_WCNC__2019__Paper__BMBM.git/IEEE_WCNC__2019__Paper__BMBM.tex
Done for now
OH LA BOULETTE : en anglais, « repartition » n'existe pas. On dit « distribution » ou « allocation »
- https://www.wordreference.com/enfr/repartition
- https://www.deepl.com/translator#en/fr/repartition

In a first scenario, we consider a total number of $N=1000$ IoT devices, with a non-uniform repartition of static devices given by $10\%,30\%,30\%,30\%$ for the four channels.
In other words, the channels are occupied\footnote{~Note that we consider higher occupancy rates that the ones considered previously in Section~\ref{sec:4:gnuradio}, because the goal of this experimental section is to evaluate our approach that uses learning to optimize the way each device retransmits its packets. We need to have a large enough probability $p_c$ of retransmission if we want to see a difference between the different learning heuristics, which is why we consider a network more densely occupied than the one considered in Section~\ref{sec:4:gnuradio}.}
respectively $10\%$, $30\%$, $30\%$, and $30\%$ of time, and the contention Markov process considered is given by $\mathrm{MaxBackOff} = 5$, and $m=5$.
In Figure~\ref{fig:43:mainExperiment1}, we show the successful transmission rate versus the number of slots, for all the proposed heuristics.

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.80\linewidth]{ResultsUCB.eps}
	\caption[First comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random, \UCB'', ``$K$ \UCB'', and ``Delayed \UCB''.]{
		Comparison between the exposed heuristics for the retransmission:`` \textcolor{blue}{Only \UCB}'', ``\textcolor{cyan}{Random}'', ``\textcolor{purple}{\UCB}'', ``\textcolor{green}{$K$ \UCB}'', and ``\textcolor{red}{Delayed \UCB}''.
		The usage of the same learning policy for transmissions and retransmission is named ``\textcolor{blue}{Only \UCB{}}'',
		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as ``no \UCB{}''.
		First scenario: learning helps but learning to retransmit smartly is not needed, as we observe that the \textcolor{cyan}{random retransmission} heuristic achieves similar performance than the others.
		We considered $N=1000$ static IoT devices, that occupy the $4$ channels in a static way leading to mean occupancy rates of $10\%,30\%,30\%,30\%$.
	}
	\label{fig:43:mainExperiment1}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.80\linewidth]{ResultsUCB2.eps}
	\caption[Second comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random'', ``\UCB'', ``$K$ \UCB'', and ``Delayed \UCB''.]{
		Second scenario: learning helps a lot (a gain of $30\%$ in terms of collision probability), and learning to retransmit smartly is needed.
		We observe that the \textcolor{cyan}{random retransmission} achieves poor performance compared to the others.
		We considered $N=2000$ static IoT devices, that occupy the $4$ channels in a static way leading to mean occupancy rates of $10\%,30\%,30\%,30\%$.
	}
	\label{fig:43:mainExperiment2}
\end{figure}

>>>>

In a first scenario, we consider a total number of $N=1000$ IoT devices, with a non-uniform distribution of static devices given by $10\%,30\%,30\%,30\%$ for the four channels.
In other words, the channels are occupied\footnote{~Note that we consider higher occupancy rates that the ones considered previously in Section~\ref{sec:4:gnuradio}, because the goal of this experimental section is to evaluate our approach that uses learning to optimize the way each device retransmits its packets. We need to have a large enough probability $p_c$ of retransmission if we want to see a difference between the different learning heuristics, which is why we consider a network more densely occupied than the one considered in Section~\ref{sec:4:gnuradio}.}
respectively $10\%$, $30\%$, $30\%$, and $30\%$ of time, and the contention Markov process considered is given by $\mathrm{MaxBackOff} = 5$, and $m=5$.
In Figure~\ref{fig:43:mainExperiment1}, we show the successful transmission rate versus the number of slots, for all the proposed heuristics.

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.90\linewidth]{ResultsUCB.eps}
	\caption[First comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random, \UCB'', ``$K$ \UCB'', and ``Delayed \UCB''.]{
		Comparison between the exposed heuristics for the retransmission:`` \textcolor{blue}{Only \UCB}'', ``\textcolor{cyan}{Random}'', ``\textcolor{purple}{\UCB}'', ``\textcolor{green}{$K$ \UCB}'', and ``\textcolor{red}{Delayed \UCB}''.
		The usage of the same learning policy for transmissions and retransmission is named ``\textcolor{blue}{Only \UCB{}}'',
		whereas the usage of a random channel selection, for both transmission and retransmission, is labeled as ``no \UCB{}''.
		First scenario: learning helps but learning to retransmit smartly is not needed, as we observe that the \textcolor{cyan}{random retransmission} heuristic achieves similar performance than the others.
		We considered $N=1000$ static IoT devices, that occupy the $4$ channels in a static way leading to mean occupancy rates of $10\%,30\%,30\%,30\%$.
	}
	\label{fig:43:mainExperiment1}
\end{figure}

@@

\begin{figure}[h!]  % [htbp]
	\centering
	\includegraphics[width=0.90\linewidth]{ResultsUCB2.eps}
	\caption[Second comparison between the exposed heuristics for the retransmission: ``Only \UCB'', ``Random'', ``\UCB'', ``$K$ \UCB'', and ``Delayed \UCB''.]{
		Second scenario: learning helps a lot (a gain of $30\%$ in terms of collision probability), and learning to retransmit smartly is needed.
		We observe that the \textcolor{cyan}{random retransmission} achieves poor performance compared to the others.
		We considered $N=2000$ static IoT devices, that occupy the $4$ channels in a static way leading to mean occupancy rates of $10\%,30\%,30\%,30\%$.
	}
	\label{fig:43:mainExperiment2}
\end{figure}
